LME-MPM applied to quasi-brittle fracture

### 2020-07-30 2020_EFM_MPM_Eigensoftening.tex
Cambio de revista

% \journal{Engineering Fracture Mechanics}
\journal{Computers and Structures}

@@

failure criterion is plotted in figure \label{fig:Damage-ft-wc}, where

@@

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist of loading
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm, supported by two rollers at a span of 47.5 mm, by
another roller on its top mid-span. Notches with a length of 16.6 mm are
created with the different notch angles of 0$^º$, 30$^º$ and 60$^º$ with respect
to the vertical axis in order to investigate the influence of the
notch angle on the peak load \ref{fig:plot-Angle-Forces-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone rock are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\ $GPa$       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ $g/cm^3$ \\
    Tensile strength (\gls{ft}) & 0.6\ $MPa$       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ $mm$ \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally, a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
was generated. The velocities of the top boundary particles are $v_x$ = 0, $v_y$ = −
2 mm/s, whereas the velocities of the bottom boundary particles are
$v_y$ = 0 mm/s. To reduce dynamic effects due to a sudden loading, a soft loading ramp has been employed as a loading curve. Finally, the model was
calibrated with a regularization parameter \gls{Ceps} = 1.7 and a
bandwidth value \gls{heps} = 15.

@@

The paper proposes the eigensoftening algorithm in a \acrshort{MPM} framework for quasi-brittle materials to overcome the limitations observed in the EigenMPM \cite{Zhang_EE_2020}. The overestimation of tensile 
stress and strain peaks in the fracture mechanism of quasi-brittle
materials and the stress oscillations observed with conventional
\acrshort{mpm} are faced. Based on the
EigenMPM approach to fracture, two additional improvements have been proposed in the present research in order to solve the shortcomings of the original concept pointed out in the aforementioned paper \cite{Zhang_EE_2020}.

@@

@@

\begin{algorithm}[H]
  \DontPrintSemicolon
  \KwInput{For each $p$, $\epsilon$-neighbourhood, $f_{t,p}$, $h_{\epsilon,p}$, $w_c$}
  \KwOutput{Return damage parameter $\chi := \{\chi_p\}$}
  % \KwData{Testing set $x$}
  $\chi_p \leftarrow \chi_p^{k}$ \;
  \For{$p$ to $N_p$}
  {
    \If{$\chi_p = 0$ \AND $\epsilon_{f,p} = 0$}
    {
      \For{$q \in B_{\epsilon,p}$}
      {
        $\sigma_{q,I} \leftarrow \text{getEigenvaluesOf}(\sigma_{q})$ \;
        \If{$\chi_q < 1$}
        {
          $\sum m_p\sigma_{p,I} \leftarrow \sum m_p\sigma_{p,I} + m_q\sigma_{q,I}$ \;
        }
        $m_p \leftarrow m_p + m_q$ \;
      }
      $\sigma_{p,\epsilon} \leftarrow \frac{1}{m_p} \sum m_p\sigma_{p,I}$\;
      \If{$\sigma_{p,\epsilon} > f_{t,p}$}
      {
        $\epsilon_{f,p} = \epsilon_{I,p}$ \;  
      }
      \ElseIf{$\chi_p \neq 1$ \AND $\epsilon_{f,p} > 0$}
      {
        $\chi_p^{k+1} \leftarrow \min\Big \{1 , \max \{\chi_p^{k},
        \frac{(\epsilon_{I,p}- \epsilon_{f,p})\ h_{\epsilon,p}}{w_c}
        \} \Big \}$ \;
      }
    }
  }
  \caption{Compute damage parameter $\chi_p^{k+1}$}
  \label{alg-eigens}
\end{algorithm}

>>>>

\journal{Engineering Fracture Mechanics}
% \journal{Computers and Structures}

@@

failure criterion is plotted in Figure \ref{fig:Damage-ft-wc}, where

@@

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist of loading
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm, supported by two rollers at a span of 47.5 mm, by
another roller on its top mid-span. Notches with a length of 16.6 mm are
created with the different notch angles of 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$ with respect
to the vertical axis in order to investigate the influence of the
notch angle on the peak load \ref{fig:plot-Angle-Forces-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone rock are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\ $GPa$       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ $g/cm^3$ \\
    Tensile strength (\gls{ft}) & 0.6\ $MPa$       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ $mm$ \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally, a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
was generated. The velocity of the top boundary particles is 
$v_y$ = − 2 mm/s, whereas the velocity of the bottom boundary particles is
$v_y$ = 0 mm/s. To reduce dynamic effects due to a sudden loading, a soft loading ramp has been employed as a loading curve. Finally, the model was
calibrated with a regularization parameter \gls{Ceps} = 1.7 and a
bandwidth value \gls{heps} = 15.

@@

The paper proposes the eigensoftening algorithm in a \acrshort{mpm} framework for quasi-brittle materials to overcome the limitations observed in the EigenMPM \cite{Zhang_EE_2020}. The overestimation of tensile 
stress and strain peaks in the fracture mechanism of quasi-brittle
materials and the stress oscillations observed with conventional
\acrshort{mpm} are faced. Based on the
EigenMPM approach to fracture, two additional improvements have been proposed in the present research in order to solve the shortcomings of the original concept pointed out in the aforementioned paper \cite{Zhang_EE_2020}.

@@

\printglossaries

@@

\begin{algorithm}[H]
  \DontPrintSemicolon
  \KwInput{For each $p$, $\epsilon$-neighbourhood, $f_{t,p}$, $h_{\epsilon,p}$, $w_c$}
  \KwOutput{Return damage parameter $\chi = \{\chi_p\}$}
  % \KwData{Testing set $x$}
  $\chi_p \leftarrow \chi_p^{k}$ \;
  \For{$p$ to $N_p$}
  {
    \If{$\chi_p = 0$ \And $\epsilon_{f,p} = 0$}
    {
      \For{$q \in B_{\epsilon,p}$}
      {
        $\sigma_{q,I} \leftarrow \text{getEigenvaluesOf} (\sigma_{q}) $ \;
        \If{$\chi_q < 1$}
        {
          $ \sum m_p\sigma_{p,I} \leftarrow \sum m_p\sigma_{p,I} + m_q\sigma_{q,I}$ \;
        }
        $ m_p \leftarrow m_p + m_q$ \;
      }
      $\sigma_{p,\epsilon} \leftarrow \frac{1}{m_p} \sum m_p\sigma_{p,I}$\;
      \If{$\sigma_{p,\epsilon} > f_{t,p}$}
      {
        $\epsilon_{f,p} = \epsilon_{I,p}$ \;  
      }
      \ElseIf{$\chi_p \neq 1$ \And $\epsilon_{f,p} > 0$}
      {
        $\chi_p^{k+1} \leftarrow \min\Big \{1 , \max \{\chi_p^{k},
        \frac{(\epsilon_{I,p}- \epsilon_{f,p})\ h_{\epsilon,p}}{w_c}
        \} \Big \}$ \;
      }
    }
  }
    \label{alg-eigens}
  \caption{Compute damage parameter $\chi$}
\end{algorithm}

### 2020-07-30 Author-Novelty.tex
Cambio de revista

\centering
\section*{Highlights}
\setlength{\parskip}{1cm plus 5mm minus 4mm}
\begin{itemize}
\item Fracture mechanics within the Material Point Method is a challenging research fields that is being improved day-to-day.

\item Eigenerosion is one of the fracture models that has been validated within dynamic problems with excellent results. The Eigensoftening algorithm is an interesting improvement of Eigenerosion for quasi-brittle materials.

\item The Local Maximum Entropy shape functions together with a Predictor-Corrector time integration scheme enhance the traditional MPM scheme.

\item Several results of the enhanced MPM with eigenerosion and eigensoftening models are provided in order to compare the performance of the proposed methodology against analytical and experimental solutions with excellent results.

>>>>

\centering
\section*{Highlights}
\setlength{\parskip}{1cm plus 5mm minus 4mm}
\begin{itemize}
\item Fracture mechanics within the MPM is a challenging research field

\item The Eigensoftening algorithm is an improvement to simulate quasi-brittle materials

\item The LME shape functions together with a NPC time integration enhance the standard MPM

\item The Eigensoftening-MPM has been validated against analytical and experimental results
